Maintenance planning on French military aircraft operations

### 2018-02-15 R/roadef_presentation/roadef_presentation.Rmd
Changes in ROADEF presentation

---
title: "ROADEF"
author: "Franco Peschiera"
date: "February 14, 2018"
output: 
  slidy_presentation:
    css: ./../../css/baobab_MH.css
    incremental: yes
  beamer_presentation:
    incremental: yes
nocite: '@*'
bibliography: MFMP.bib
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
setwd("../")
source('scripts/params.R')
source('scripts/import_results.R')
exp_directory = PATHS[['experiments']] %>% paste0('201802141539/')
gantt <- print_solution(exp_directory, max_resources = 30)
```

@@

@@

<div style="font-size: 20px;">
```{r}
gantt
```
</div>

@@

A series of $j \in \mathcal{J}$ tasks are planned along a horizon divided into $t \in \mathcal{T}$ periods. Since all tasks are already scheduled, we know time periods $T_j \subset \mathcal{T}$ in which they will be realized. Similarly, all the tasks to be realized in period $t$ are known and defined by set $J_t \subset \mathcal{J}$.

Each task requires a certain number $r_{j}$ of resources $i \in \mathcal{I}$ which it employs for a time duration defined by $h_j$ in each period. Set $a_{j} \subset \mathcal{I}$ lists the resources that can be assigned to each task and set $O_i \subset \mathcal{J}$ consists of tasks for which resource $i$ can be used. The number of resources in use for each period $d_t$ is also known.

@@

The problem was born out of the need of the french Air Force to maximize the availability of their Mirage 2000 fleet.

@@

* In @Cho2011, US Army aircraft was assigned daily operations over a year to aircraft.
* In @Kozanidis2008, Greek aircraft had monthly assignments of maintenances and flight hours. 
* In @Verhoeff2015, monthly assignments were also done, and several objectives were taken into account: availability, serviceability and sustainability.
* In @Chastellux2017, a first version of the model was done, assigning missions instead of flight hours.

@@

The present model assigns missions to aircraft, permitting to add more complex functionality, such as minimum assignment durations.

This model also features several improvements in the MIP modelling that permit greater instance sizes.

@@

|Set|Description|
|-----|--------------------------------------------------------------------|
|    $\mathcal{T}_j$     |  time periods $t \in \mathcal{T}$ in which task $j$ is active. |
|    $\mathcal{J}_t$    |  tasks $j \in \mathcal{J}$ to be realized in period $t$. |
|    $\mathcal{I}_j$     |  resources $i \in \mathcal{I}$ that can be assigned to task $j$. |
|    $\mathcal{O}_i$     |  tasks $j \in \mathcal{J}$ for which resource $i$ can be used. |
|    $\mathcal{T}^{s}_t$ |  time periods $t' \in \mathcal{T}$ such that $t' \in \{ \max{\{1, t - M+1\}},  ..., {t}$\}. |

@@

Only *main parameters* are shown

@@

<!-- ## Model: parameters (2) -->

<!-- |Parameters|Description| -->
<!-- |--------------|----------------------------------------------------------------------| -->
<!-- | $N_t$              | number of resources in already-planned maintenances in period $t$ at the beginning | of the planning ho |izon.| -->
<!-- | $D_t$              | number of resources to be assigned in total in period $t$. | -->
<!-- | $Rut^{Init}_{i}$   | remaining usage time for resource $i$ at the start of the planning horizon. | -->
<!-- | $Ret^{Init}_{i}$   | remaining elapsed time for resource $i$ at the start of the planning horizon. | -->
<!-- | $Ret^{Init}_{sum}$ | sum of remaining elapsed times at the start of the planning horizon. | -->
<!-- | $Rut^{Init}_{sum}$ | sum of remaining elapsed time at the start of the planning horizon. | -->

@@

$$
\begin{align}
&\sum_{t' \in \mathcal{T}^{s}_t} \sum_{i \in \mathcal{I}} m_{it'} + N_t \leq m_{max}
&t \in \mathcal{T} \\
&\sum_{t' \in \mathcal{T}^{s}_t} \sum_{i \in \mathcal{I}} m_{it'} + N_t + D_t\leq u_{max} 
&t \in \mathcal{T}\\
&\sum_{i \in \mathcal{I}_j} a_{jti} = R_j
&j \in \mathcal{J}, t \in \mathcal{T}_j\\
&\sum_{t' \in \mathcal{T}^{s}_t} m_{it'} + \sum_{j \in \mathcal{J}_t \cap \mathcal{O}_i} a_{jti} \leq 1 
& t \in \mathcal{T}, i \in \mathcal{I}
\end{align}
$$

@@

Only usage time is shown. Elapsed times are treated in the same way.

$$
\begin{align}
& rut_{it} \leq rut_{it-1} + H m_{it} - \sum_{j \in \mathcal{J}_t \cap \mathcal{O}_i} a_{jti} H_j & t =1, ..., \mathcal{T}, i \in \mathcal{I}\\
& rut_{i0} = Rut^{Init}_i
        & i \in \mathcal{I}\\
& rut_{it} \geq H m_{it}
        & t \in \mathcal{T}, i \in \mathcal{I}\\ 
& rut_{it} \in [0,H]
        & t \in \mathcal{T}, i \in \mathcal{I} \\
& \sum_{i \in \mathcal{I}} rut_{it} \geq Rut^{Init}_{sum}
        & t = |\mathcal{T}|
\end{align}
$$

@@

## Multi-objective

@@

>>>>

---
title: "Maintenance planning on French military aircraft operations"
author: "Franco Peschiera, Alain Haït, Olga Battaïa, Nicolas Dupin"
date: "February 14, 2018"
output: 
  beamer_presentation:
    theme: "Madrid"
    toc: true
    slide_level: 2
    incremental: yes
    fig_height: 4
  slidy_presentation:
    css: ./../../css/baobab_MH.css
    incremental: yes
    footer: "ISAE-SUPAERO, Université de Toulouse, France"
nocite: '@*'
bibliography: MFMP.bib
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
setwd("../")
source('scripts/params.R')
source('scripts/import_results.R')
# library(png)
# library(grid)
exp_directory = PATHS[['experiments']] %>% paste0('201802141539/')
gantt <- print_solution(exp_directory, max_resources = 30)
show_gantt <- knitr::opts_knit$get("rmarkdown.pandoc.to")=="slidy"
```

@@

Flight and Maintenance Planning (FMP) problems are raised in civilian or military operational applications.

They schedule maintenances of aircraft taking into account the implications of maintenance immobilizations for the demand in flight hours, for a conjoint optimization among maintenance dates and flight hours.

@@

```{r, echo=FALSE, out.width='100%'}
if (show_gantt){
    htmltools::HTML('<div style="font-size: 20px;width:100%">')
    gantt    
    htmltools::HTML("</div>")
} else {
    knitr::include_graphics('./../../img/calendar.png')
}

```

@@

* A series of $j \in \mathcal{J}$ tasks are planned along a horizon divided into $t \in \mathcal{T}$ periods. Since all tasks are already scheduled, we know time periods $T_j \subset \mathcal{T}$ in which they will be realized. Similarly, all the tasks to be realized in period $t$ are known and defined by set $J_t \subset \mathcal{J}$. 
* Each task requires a certain number $r_{j}$ of resources $i \in \mathcal{I}$ which it employs for a time duration defined by $h_j$ in each period. Set $a_{j} \subset \mathcal{I}$ lists the resources that can be assigned to each task and set $O_i \subset \mathcal{J}$ consists of tasks for which resource $i$ can be used. The number of resources in use for each period $d_t$ is also known.

@@

This thesis was started out of the need of the french Air Force to maximize the availability of their Mirage 2000 fleet while controlling cost and maintenance capacity.

@@

* In @Cho2011, US Army aircraft were assigned daily operations over a year to aircraft.
* In @Kozanidis2008, Greek aircraft had monthly assignments of maintenances and flight hours. 
* In @Verhoeff2015, monthly assignments were done and several objectives were taken into account: availability, serviceability and sustainability.
* In @Chastellux2017, an initial version of the model was done, assigning missions instead of only flight hours.

@@

* The present model deals with an heterogeneus fleet, with different standards and capacities.
* Multiple objectives have been incorporated in order to satisfy the French Air Force's specific needs.
* This model also features several improvements in the MIP modelling that permit greater instance sizes.

@@

|Set|Description|
|-----|--------------------------------------------------------------------|
|    $\mathcal{I}$       |  resources. |
|    $\mathcal{T}$       |  time periods inside planning horizon. |
|    $\mathcal{J}$       |  tasks. |
|    $\mathcal{T}_j$     |  time periods $t \in \mathcal{T}$ in which task $j$ is active. |
|    $\mathcal{J}_t$    |  tasks $j \in \mathcal{J}$ to be realized in period $t$. |
|    $\mathcal{I}_j$     |  resources $i \in \mathcal{I}$ that can be assigned to task $j$. |
|    $\mathcal{O}_i$     |  tasks $j \in \mathcal{J}$ for which resource $i$ can be used. |
|    $\mathcal{T}^{s}_t$ |  time periods $t' \in \mathcal{T}$ such that $t' \in \{ \max{\{1, t - M+1\}},  ..., {t}$\}. |

@@

(Only **main parameters** are shown).

@@

## Model: parameters (2)

(**secondary parameters**).

|Parameters|Description|
|--------------|----------------------------------------------------------------------|
| $N_t$              | number of resources in already-planned maintenances in period $t$ at the beginning | of the planning ho |izon.|
| $D_t$              | number of resources to be assigned in total in period $t$. |
| $Rut^{Init}_{i}$   | remaining usage time for resource $i$ at the start of the planning horizon. |
| $Ret^{Init}_{i}$   | remaining elapsed time for resource $i$ at the start of the planning horizon. |
| $Ret^{Init}_{sum}$ | sum of remaining elapsed times at the start of the planning horizon. |
| $Rut^{Init}_{sum}$ | sum of remaining elapsed time at the start of the planning horizon. |

@@

<div class="incremental">

The two components of the objective function: the max number of maintenance and the max number of unavailable aircraft.
\begin{align}
    &\sum_{t' \in \mathcal{T}^{s}_t} \sum_{i \in \mathcal{I}} m_{it'} + N_t \leq m_{max}
    &t \in \mathcal{T} \\
    &\sum_{t' \in \mathcal{T}^{s}_t} \sum_{i \in \mathcal{I}} m_{it'} + N_t + D_t\leq u_{max}
    &t \in \mathcal{T}
\end{align}

Mission's needs and incompatibility of several tasks or maintenance in the same period.
\begin{align}
&\sum_{i \in \mathcal{I}_j} a_{jti} = R_j
&j \in \mathcal{J}, t \in \mathcal{T}_j\\
&\sum_{t' \in \mathcal{T}^{s}_t} m_{it'} + \sum_{j \in \mathcal{J}_t \cap \mathcal{O}_i} a_{jti} \leq 1
& t \in \mathcal{T}, i \in \mathcal{I}
\end{align}

</div>

@@

Time usage of an aircraft depending on the assignment to a task.
<!-- <div class="incremental" style="margin-left: 4em; position: relative">  -->
<!--   <div style="position: absolute; left: 0; top: 0;background-color:white;width:100%"> -->
\begin{align}
& rut_{it} \leq rut_{it-1} + H m_{it} - \sum_{j \in \mathcal{J}_t \cap \mathcal{O}_i} a_{jti} H_j & t =1, ..., \mathcal{T}, i \in \mathcal{I}\\
& rut_{i0} = Rut^{Init}_i
        & i \in \mathcal{I}\\
& rut_{it} \geq H m_{it}
        & t \in \mathcal{T}, i \in \mathcal{I}\\
& rut_{it} \in [0,H]
        & t \in \mathcal{T}, i \in \mathcal{I} \\
& \sum_{i \in \mathcal{I}} rut_{it} \geq Rut^{Init}_{sum}
        & t = |\mathcal{T}|
\end{align}

  <!-- </div>  -->
  <!-- <div style="position: absolute; left: 0; top: 0;background-color:white;width:100%"> -->

## Model: constraints (3)

Elapsed times are treated in the same way but consumption is always 1.
\begin{align}
& ret_{it} \leq ret_{it-1} + E m_{it} - 1 & t =1, ..., \mathcal{T}, i \in \mathcal{I}\\
& ret_{i0} = Ret^{Init}_i
        & i \in \mathcal{I}\\
& ret_{it} \geq E m_{it}
        & t \in \mathcal{T}, i \in \mathcal{I}\\
& ret_{it} \in [0,E]
        & t \in \mathcal{T}, i \in \mathcal{I} \\
& \sum_{i \in \mathcal{I}} ret_{it} \geq Ret^{Init}_{sum}
        & t = |\mathcal{T}|
\end{align}

## Model: constraints (4)

Visualization of each constraints for one resource.

```{r out.width="250px", echo=FALSE}
# img <- readPNG("./../../img/balance2.png")
#  grid.raster(img)
knitr::include_graphics("./../../img/balance2.png")
```

@@

The following instances were tested:

| id | $\|\mathcal{J}\|$ |$\|\mathcal{T}\|$ |  assign |  objective  |  time (s) |   bound |
|-----|------------------|------------------|---------| ------------| -----------| --------|
| I_0 |                9 |               11 |     310 |        62.0 |        0.7 |    62.0 |
| I_1 |                9 |               21 |     650 |        63.0 |       68.7 |    63.0 |
| I_2 |                9 |               31 |     990 |        63.0 |     3600.1 |    62.0 |
| I_3 |                9 |               41 |    1249 |        64.0 |     3603.9 |    61.7 |
| I_4 |               10 |               11 |     530 |        82.0 |        0.9 |    82.0 |
| I_5 |               10 |               21 |    1070 |        83.0 |      144.0 |    83.0 |
| I_6 |               10 |               31 |    1610 |        83.0 |     3600.1 |    82.0 |
| I_7 |               10 |               41 |    2069 |        84.0 |     3609.1 |    81.8 |
| I_8 |               11 |               11 |    1080 |       139.0 |      530.6 |   139.0 |
| I_9 |               11 |               21 |    2120 |       149.0 |     3600.0 |   139.9 |

## Multi-objective (1)

Comparison of two objectives: max availability or maintenance capacity.
Different weights were used for the two objectives: $W_1, W_2 \in [0..1]$ with pace of 0.1
```{r out.width="200pt", echo=FALSE}
# img <- readPNG()
#  grid.raster(img)
knitr::include_graphics("./../../img/multiobjective.png")

```

Pareto diagram for instance id=I_5.

## Multi-objective (2)

* Analysis for all instances: not many pareto optimal points available.
* Between 2 and 4 for all instances
* The objectives are quite aligned one with the other.

@@

* **Try different objectives**: availability and maintenances.
* **Try more linent objectives**: relax the `max` in the objective.
* **Break symmetries**: clustering candidates, fixing end availabilities.
* **Add more constraints**: minimal duration of assignments, storage of aircraft.
